Linear instability underlying synchronous waving of marine grass in flow

### 2014-09-06 Paper/grass.tex
Abstract changed
Included consistent notation for "monami"

\graphicspath{ {images/}{images/Growth_Rate_Cont_Plots/} }
\newcommand{\bn}{{\boldsymbol{\hat{n}}}}
\newcommand{\bt}{{\boldsymbol{\hat{t}}}}
\newcommand{\bu}{\mathbf{u}}
\newcommand{\grad}{\mathbf{\nabla}}
\newcommand{\del}{\partial}
\newcommand{\hg}{h_g}
\newcommand{\Rey}{\text{R}}
\newcommand{\Ndg}{\tilde{N}_g}
\newcommand{\monami}{\textit{monami}}
% \newcommand{\shreyas}[1]{{\bf (#1)}}
\newcommand{\shreyas}[1]{}
\definecolor{tableShade}{gray}{0.8}

@@

%The spontaneous waving of marine grass is thought to be due to a Kelvin-Helmholtz instability resulting from an inflection point near the canopy top in the flow profile.
We study the mechanism behind \monami, a phenomenon associated with the synchronous waving of marine grass. Linear stability analysis of a mathematical model incorporating grass 
drag as a continuum field has been 
carried out. Our analysis predicts a critical threshold condition for flow instability leading to \monami and associated waving frequency
which compares well with the experimental observations. Further analysis predicts the existence of two distinct modes of instability which we refer as ``Mode 1'' and ``Mode 2''.
%Mode 1 has many characteristics similar to the Kelvin-Helmholtz instability where shear layer formed near grass tip due to grass drag discontinuity plays a critical role. 
%Mode 2 can be interpreted as instability of 
%inviscid flow through a grass bed where shear layer doesn't play an important role. 
The occurrence of these modes for flow instability depends on submergence ratio and the grass 
number density. Asymptotic analysis along with numerical computation in the limit of dense grass bed has been carried out. The analysis reveals that Mode 1 represents flow instability on the 
scale of shear layer which is formed near grass tip due to drag discontinuity near the canopy top, whereas Mode 2 can be understood as instability on the scale of water column. 
For the parameter regime typical of marine grass beds, the two modes merge and are indistinguishable.

@@

\end{abstract}
\maketitle
%\section{Introduction}
Sea grasses exhibit rich set of dynamics due to their interaction with the flow of water and can considerably affect surrounding hydrodynamic conditions.
Resulting changes in hydrodynamic conditions can influence number of environmental variables and processes such as transport of sediments, contaminants, dissolved oxygen, plant growth, and biomass production, etc \cite{Fonseca87,Nepf99,Nepf2012}. 
Flexible grass canopies respond to steady currents in the form of coherent large amplitude oscillations, known as \monami \cite{AckermanOkubo93}. Such motion is considered to play a central role in the recruitment 
of microscopic marine organisms, such as blue mussel larvae\cite{Grizzle96}.
The hydrodynamic mechanism underlying \monami is the focus of this paper.

>>>>

\graphicspath{ {images/}{images/Growth_Rate_Cont_Plots/} }
\newcommand{\bn}{{\boldsymbol{\hat{n}}}}
\newcommand{\bt}{{\boldsymbol{\hat{t}}}}
\newcommand{\bu}{\mathbf{u}}
\newcommand{\grad}{\mathbf{\nabla}}
\newcommand{\del}{\partial}
\newcommand{\hg}{h_g}
\newcommand{\Rey}{\text{R}}
\newcommand{\Ndg}{\tilde{N}_g}
\newcommand{\monami}{\textit{monami }}
% \newcommand{\shreyas}[1]{{\bf (#1)}}
\newcommand{\shreyas}[1]{}
\definecolor{tableShade}{gray}{0.8}

@@

%The spontaneous waving of marine grass is thought to be due to a Kelvin-Helmholtz instability resulting from an inflection point near the canopy top in the flow profile.
%We study the mechanism behind \monami, a phenomenon associated with the synchronous waving of marine grass. Linear stability analysis of a mathematical model incorporating grass 
%drag as a continuum field has been 
%carried out. Our analysis predicts a critical threshold condition for flow instability leading to \monami and associated waving frequency
%which compares well with the experimental observations. Further analysis predicts the existence of two distinct modes of instability which we refer as ``Mode 1'' and ``Mode 2''.
%Mode 1 has many characteristics similar to the Kelvin-Helmholtz instability where shear layer formed near grass tip due to grass drag discontinuity plays a critical role. 
%Mode 2 can be interpreted as instability of 
%inviscid flow through a grass bed where shear layer doesn't play an important role. 
%The occurrence of these modes for flow instability depends on submergence ratio and the grass 
%number density. Asymptotic analysis along with numerical computation in the limit of dense grass bed has been carried out. The analysis reveals that Mode 1 represents flow instability on the 
%scale of shear layer which is formed near grass tip due to drag discontinuity near the canopy top, whereas Mode 2 can be understood as instability on the scale of water column. 
%For the parameter regime typical of marine grass beds, the two modes merge and are indistinguishable.

@@

We show that the threshold flow condition for the spontaneous waving of marine grass arises due to the drag exerted by the vegetation on the flow. We show this by developing a self-consistent model of the flow past rigid vegetation. Linear stability analysis of the steady flow resulting from this model naturally leads to the threshold condition, and as well as the origin of the mechanistic threshold. Asymptotic analysis of the model for dense grass reveals two unstable modes. One of these modes shares many common features with the Kelvin-Helmholtz instability, but is demonstrated to be different from it. The predictions of the model for threshold condition and the resulting dominant frequency of waving agree with laboratory observations. 
\end{abstract}
\maketitle
%\section{Introduction}
Sea grasses exhibit rich set of dynamics due to their interaction with the flow of water and can considerably affect surrounding hydrodynamic conditions.
Resulting changes in hydrodynamic conditions can influence number of environmental variables and processes such as transport of sediments, contaminants, dissolved oxygen, plant growth, and biomass production, etc \cite{Fonseca87,Nepf99,Nepf2012}. 
Flexible grass canopies respond to steady currents in the form of coherent large amplitude oscillations, known as \monami \cite{AckermanOkubo93}. Such motion is considered to play a central role in the recruitment 
of microscopic marine organisms, such as blue mussel larvae\cite{Grizzle96}.
The hydrodynamic mechanism underlying \monami is the focus of this paper.
